SHARP - II. Mass structure in strong lenses is not necessarily dark matter substructure: A flux ratio anomaly from an edge-on disc

### 2016-03-13 papers/B1555_EdgeOn/B1555_letter.tex
enclose_mass [unit check] half way

\begin{table*}
\centering
\caption{A summary of the observational constraints and the predicted lens-model parameters. The lensed image positions in the radio \citep{Marlow99} (in arcsec) and average flux ratios \citep{K03} are with respect to image A. Uncertainties in the position measurements for images A, B and C are $0.001$~arcsec and for image D are $0.006$~arcsec.}
\begin{tabular}{cccccccc}
\hline
Image	&\multicolumn{3}{c}{Observed MERLIN} 	 	& \multicolumn{4}{c}{Model-Predicted}\\
		&\multicolumn{3}{c}{5 GHz}		& & & {SIE+ExpDisc} & SIE-only\\
		&East &North & Flux ratio &East 	&North & Flux ratio &Flux ratio\\ 
\hline
A  &$0$    		&$0$			&  $1$ 				&$+0.000$  &$+0.000$	& $1$ 		& $1$\\  
B  &$-0.0726$	&$+0.0480$	& $0.620 \pm 0.039$ 	&$-0.0715$ &$+0.0487$		& $0.606$ 	& $0.971$ \\  
C  &$-0.4117$ 	&$-0.0280$	& $0.507\pm 0.030$	&$-0.4115$ &$-0.0278$		& $0.445$ 	& $0.312$\\  
D  &$-0.1619$	&$-0.3680$	& $0.086 \pm 0.024$ 	&$-0.1693$ &$-0.3704$		& $0.123$ 	& $0.106$\\  
\hline
\end{tabular}
\label{tab:results}
\end{table*}

\begin{table}
\centering
\caption{Best-fit model parameters and uncertainties for our two-component SIE + exponential disc lens model (our single SIE model is effectively equivalent to the model in \citet{Marlow99} and so is not shown here). The positions are measured in arcsec, relative to the MERLIN radio position of image A. The ellipticity is defined as $e=1-q$ and the position angles are measured in degrees east of north. The Einstein radius $b$ and disc scale radius are measured in arcsec.}
\begin{tabular}{ccc}
\hline 
Parameter    & SIE Component & ExpDisc Component  \\
%   & (SIE) & (ExpDisc)\\
\hline
$\Delta$RA	& $-0.177 \pm 0.003$		& $-0.162$ \textit{(fixed)}\\
$\Delta$Dec	& $-0.205 \pm 0.004$		& $-0.206$ \textit{(fixed)} \\
$b$ 			& $0.178 \pm 0.004$  		& ...  \\
$\kappa_0$ & ... &$0.273 \pm 0.003$\\
$e$	  		& $0.25 \pm 0.05$			& $0.83$ \textit{(fixed)} \\
$\theta$ 		& $4.9 \pm 1.3$			& $8.0$ \textit{(fixed)}	 \\
$R_s$			& ...  						& $0.236 \pm 0.003$	 \\
\hline
\label{tab:model}
\end{tabular}
\end{table}

>>>>

\begin{table*}
\centering
\caption{A summary of the observational constraints and the predicted lens-model parameters. The lensed image positions in the radio \citep{Marlow99} (in arcsec) and average flux ratios \citep{K03} are with respect to image A. Uncertainties in the position measurements for images A, B and C are $0.001$~arcsec and for image D are $0.006$~arcsec.}
\begin{tabular}{cccccccc}
\hline
Image	&\multicolumn{3}{c}{Observed} 	 	& \multicolumn{4}{c}{Model-Predicted}\\
		&\multicolumn{3}{c}{MERLIN 5 GHz}		& & & {SIE+ExpDisc} & SIE-only\\
		&East &North & Flux ratio &East 	&North & Flux ratio &Flux ratio\\ 
\hline
A  &$0$    		&$0$			&  $1$ 				&$+0.000$  &$+0.000$	& $1$ 		& $1$\\  
B  &$-0.0726$	&$+0.0480$	& $0.620 \pm 0.039$ 	&$-0.0715$ &$+0.0487$		& $0.606$ 	& $0.971$ \\  
C  &$-0.4117$ 	&$-0.0280$	& $0.507\pm 0.030$	&$-0.4115$ &$-0.0278$		& $0.445$ 	& $0.312$\\  
D  &$-0.1619$	&$-0.3680$	& $0.086 \pm 0.024$ 	&$-0.1693$ &$-0.3704$		& $0.123$ 	& $0.106$\\  
\hline
\end{tabular}
\label{tab:results}
\end{table*}

\begin{table}
\centering
\caption{Best-fit model parameters and uncertainties for our two-component SIE + exponential disc lens model (our single SIE model is effectively equivalent to the model in \citet{Marlow99} and so is not shown here). The positions are measured in arcsec, relative to the MERLIN radio position of image A. The ellipticity is defined as $e=1-q$ and the position angles are measured in degrees east of north. The Einstein radius $b$ and disc scale radius ($R_s$) are measured in arcsec.}
\begin{tabular}{ccc}
\hline 
Parameter    & SIE Component & ExpDisc Component  \\
%   & (SIE) & (ExpDisc)\\
\hline
$\Delta$RA	& $-0.177 \pm 0.003$		& $-0.162$ \textit{(fixed)}\\
$\Delta$Dec	& $-0.205 \pm 0.004$		& $-0.206$ \textit{(fixed)} \\
$b$ 			& $0.178 \pm 0.004$  		& ...  \\
$\kappa_0$ & ... &$0.273 \pm 0.003$\\
$e$	  		& $0.25 \pm 0.05$			& $0.83$ \textit{(fixed)} \\
$\theta$ 		& $4.9 \pm 1.3$			& $8.0$ \textit{(fixed)}	 \\
$R_s$			& ...  						& $0.236 \pm 0.003$	 \\
\hline
\label{tab:model}
\end{tabular}
\end{table}

### 2016-03-13 papers/B1555_EdgeOn/arXiv_v2/B1555_letter.tex
enclose_mass [unit check] half way

In this letter, we explore the idea that not all flux-ratio anomalies are due to the presence of clumpy substructure, but that some originate from more complex lens galaxy mass distributions than have initially been considered. To this end, we use new infrared \textbf{and radio interferometric} imaging data from the Strong lensing at High Angular Resolution Programme \citep[SHARP;][Chen et al. 2016 submitted]{mckean07,lagattuta10,SHARP12, V12}. The main goal of SHARP is to investigate galaxy structure and other astrophysical topics through the re-observation of known gravitational lens systems at high angular resolution, making use of observations with (1) adaptive optics (AO) on the W. M. Keck telescope, (2) the {\it Hubble Space Telescope (HST)}, and (3) large radio and mm interferometric arrays.

@@

There are several notable features in the high-resolution AO and {\it HST} data. These include the nearly complete lack of emission associated with lensed images B and D, 
%as can be seen by comparing the radio contours to the $K^\prime$-band emission 
(Fig.~\ref{fig:3panel}b), and the faint, but clearly visible emission from the lensing galaxy that appears to be an edge-on disc, with a position angle of $\sim 10$~deg and an axis ratio $q \sim 0.2$.  We note that the position angle of the lensing galaxy is such that the disc emission lies close to the lensed images B and D. Therefore, the lack of emission from these \textbf{rest-frame optical} images is likely due to dust extinction and additional demagnification due to lensing.

@@

\textbf{We use {\sc gravlens} to obtain the best-fit SIE+ExpDisc lens model. The best-fit parameters from \citet{Marlow99} are chosen as the starting point of the SIE profile. We have used the NICMOS F160W imaging, in which the disc component has the highest signal-to-noise ratio amongst our data, to fix the disc centroid position relative to image A to $(-0.162, -0.206)$, its ellipticity to $e=0.83$, and its position angle to $\theta=8^\circ$. The total number of free parameters is 9. After obtaining the best-fit lens model, we run a Markov-Chain Monte-Carlo (MCMC) analysis to obtain the uncertainties.}

The best-fit model parameters and 68 per cent confidence level (CL) errors are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:3panel}c. Table \ref{tab:results} also shows the comparison between the observed radio data and the model-predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth SIE model there is significant improvement in the predicted flux ratios when compared to the results for the SIE-only model. 
% with a reduced $\chi^2 = 2.7$
The positions are all in agreement, and the flux ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The remaining differences in the fluxes may be due to the real lensed source being slightly extended rather than the point source that we used in our model. The model requires that the \textbf{ExpDisc component makes up about 15 per cent of the total projected mass within the Einstein radius (1.7~kpc for an assumed lens redshift of $z_l = 1$ but the mass fraction is independent of the lens and source redshifts since this ratio is independent of distance).  In this central region of an edge-on spiral galaxy, the baryonic mass can be a significant component of the total projected mass. Investigations of edge-on spiral lenses, for example from the SWELLS survey, show that the disc mass fraction within the Einstein radius ranges from 5 to 30 per cent\citep{swells1,swells3,swells6}.} 
%In the NIR imaging there is no sign of a bulge; a bulge plus disk fit to the lens galaxy surface brightness distribution gives a bulge component that is consistent with zero emission.  We therefore interpret the ExpDisc component as a good description of the baryonic mass of the lens galaxy. 
%We note that there is a significant 0.036~arcsec offset between the SIE and ExpDisc components, which is likely due to an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models require the stronger constraints provided by deep imaging of the full extent of the gravitational arc in future high resolution radio observations.

@@

Here, we find evidence for an edge-on disc in the flux-ratio anomaly
gravitational lens B1555+375 that can account for the required
perturbation in the mass model.  \textbf{Although a constraint on the disc mass from photometry is currently not available due to the low signal-to-noise ratio in the AO imaging and the lack of redshift information, our model indicates that the flux-ratio anomaly in B1555+375 has a strong link to the presence of an edge-on disc. } Such
high-order deviations from a smooth mass distribution have already
been suggested previously \citep{evans03, congdon05}, but this is the
first time that an observed disc has been explicitly modelled to
account for a flux-ratio anomaly. However, it is important to realise
that by including observationally-motivated modifications to a
`smooth' mass distribution, such as luminous dwarf companions or
edge-on discs, in the mass models, we are still limited in our
interpretation if we can not determine how much of the anomaly are
due to these components. For our current analysis of B1555+375, we
have shown that the entire anomaly could be caused by the edge-on
disc. 
\textbf{That is not to imply this is a unique model, but rather that the
disc component is a plausible explanation of the anomaly. }
 %We note that the SWELLS system J0820+4847 has an edge-on disc that directly crosses but does not significantly distort the lensed arc emission in the system \citep{swells1}. The disc in this system has been shown to comprise 15 per cent of the mass within the Einstein radius, comparable to our results on B1555+375 and is also fully consistent with the 5 to 30 per cent contribution of discs in the other lenses in the  SWELLS sample \citep{swells1,swells3,swells6}.}

@@

Our analysis of B1555+375 demonstrates that the choice of the macromodel used to describe the `smooth' lens galaxy mass distribution, combined with the limited information provided by the position of compact lensed images, can lead to a misinterpretation of the origin of flux-ratio anomalies in some systems. \textbf{Special care may be needed for small-separation lens systems, in which the lensing galaxy is likely to be a spiral \citep[e.g.,][]{Turner84, F92}.  This is because a disc with a high inclination angle can significantly affect the lensing cross-section for such systems \citep{Maller97,Kee98,Moller03}. The current sample of lensed quasars with radio-wavelength flux ratio measurements is small.  Therefore, having even a few systems with edge-on discs may be sufficient to introduce a bias into the substructure analysis. In fact, our near-infrared AO imaging of
other lenses in the SHARP sample shows that B1555+375 is not the only
system that has a significant disc component that crosses the merging lensed images (Fassnacht et al., in prep.).  
These gravitational lenses will be investigated in future
work, but the images alone suggest that other systems may have a similar
link between flux-ratio anomalies and disc components. This potential bias introduced by edge-on discs may be avoided by post-selection as  sample sizes grow in the future. Therefore, it is critical to obtain high-resolution imaging lenses to detect additional mass features such as edge-on discs or even massive luminous satellites.} 
% A number of lens systems with significant disc components arealready known, including, for example, the sample of 16 edge-on disks from SDSS found in the SWELLS survey \citep{swells1}. 
 This result is timely, as upcoming large-scale surveys (e.g., the Dark Energy Survey, Kilo-Degree Survey, \textit{Euclid}, and the Large Synoptic Survey Telescope) will deliver orders of magnitude more gravitationally lensed quasars for which flux ratios can
  be measured using narrow lines \citep[e.g.][]{N14}.
  %Finally, we note that massive edge-on discs are \textit{not} observed in some flux-ratio anomaly systems \citep[and the presence of a disc-like component may not resolve the anomaly anyway;][]{evans03}, indicating that these systems are still useful for probing the dark matter substructure mass function.

>>>>

In this letter, we explore the idea that not all flux-ratio anomalies are due to the presence of clumpy substructure, but that some originate from more complex lens galaxy mass distributions than have initially been considered. To this end, we use new infrared and radio interferometric imaging data from the Strong lensing at High Angular Resolution Programme \citep[SHARP;][Chen et al. 2016 submitted]{mckean07,lagattuta10,SHARP12, V12}. The main goal of SHARP is to investigate galaxy structure and other astrophysical topics through the re-observation of known gravitational lens systems at high angular resolution, making use of observations with (1) adaptive optics (AO) on the W. M. Keck telescope, (2) the {\it Hubble Space Telescope (HST)}, and (3) large radio and mm interferometric arrays.

@@

There are several notable features in the high-resolution AO and {\it HST} data. These include the nearly complete lack of emission associated with lensed images B and D, 
%as can be seen by comparing the radio contours to the $K^\prime$-band emission 
(Fig.~\ref{fig:3panel}b), and the faint, but clearly visible emission from the lensing galaxy that appears to be an edge-on disc, with a position angle of $\sim 10$~deg and an axis ratio $q \sim 0.2$.  We note that the position angle of the lensing galaxy is such that the disc emission lies close to the lensed images B and D. Therefore, the lack of emission from these rest-frame optical images is likely due to dust extinction and additional demagnification due to lensing.

@@

We use {\sc gravlens} to obtain the best-fit SIE+ExpDisc lens model. The best-fit parameters from \citet{Marlow99} are chosen as the starting point of the SIE profile. We have used the NICMOS F160W imaging, in which the disc component has the highest signal-to-noise ratio amongst our data, to fix the disc centroid position relative to image A to $(-0.162, -0.206)$, its ellipticity to $e=0.83$, and its position angle to $\theta=8^\circ$. The total number of free parameters is 9. After obtaining the best-fit lens model, we run a Markov-Chain Monte-Carlo (MCMC) analysis to obtain the uncertainties.

The best-fit model parameters and 68 per cent confidence level (CL) errors are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:3panel}c. Table \ref{tab:results} also shows the comparison between the observed radio data and the model-predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth SIE model there is significant improvement in the predicted flux ratios when compared to the results for the SIE-only model. 
% with a reduced $\chi^2 = 2.7$
The positions are all in agreement, and the flux ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The remaining differences in the fluxes may be due to the real lensed source being slightly extended rather than the point source that we used in our model. The model requires that the ExpDisc component makes up about 15 per cent of the total projected mass within the Einstein radius (1.7~kpc for an assumed lens redshift of $z_l = 1$ but the mass fraction is independent of the lens and source redshifts since this ratio is independent of distance).  In this central region of an edge-on spiral galaxy, the baryonic mass can be a significant component of the total projected mass. Investigations of edge-on spiral lenses, for example from the SWELLS survey, show that the disc mass fraction within the Einstein radius ranges from 5 to 30 per cent\citep{swells1,swells3,swells6}.
%In the NIR imaging there is no sign of a bulge; a bulge plus disk fit to the lens galaxy surface brightness distribution gives a bulge component that is consistent with zero emission.  We therefore interpret the ExpDisc component as a good description of the baryonic mass of the lens galaxy. 
%We note that there is a significant 0.036~arcsec offset between the SIE and ExpDisc components, which is likely due to an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models require the stronger constraints provided by deep imaging of the full extent of the gravitational arc in future high resolution radio observations.

@@

Here, we find evidence for an edge-on disc in the flux-ratio anomaly
gravitational lens B1555+375 that can account for the required
perturbation in the mass model.  Although a constraint on the disc mass from photometry is currently not available due to the low signal-to-noise ratio in the AO imaging and the lack of redshift information, our model indicates that the flux-ratio anomaly in B1555+375 has a strong link to the presence of an edge-on disc.  Such
high-order deviations from a smooth mass distribution have already
been suggested previously \citep{evans03, congdon05}, but this is the
first time that an observed disc has been explicitly modelled to
account for a flux-ratio anomaly. However, it is important to realise
that by including observationally-motivated modifications to a
`smooth' mass distribution, such as luminous dwarf companions or
edge-on discs, in the mass models, we are still limited in our
interpretation if we can not determine how much of the anomaly are
due to these components. For our current analysis of B1555+375, we
have shown that the entire anomaly could be caused by the edge-on
disc. 
That is not to imply this is a unique model, but rather that the
disc component is a plausible explanation of the anomaly. 
 %We note that the SWELLS system J0820+4847 has an edge-on disc that directly crosses but does not significantly distort the lensed arc emission in the system \citep{swells1}. The disc in this system has been shown to comprise 15 per cent of the mass within the Einstein radius, comparable to our results on B1555+375 and is also fully consistent with the 5 to 30 per cent contribution of discs in the other lenses in the  SWELLS sample \citep{swells1,swells3,swells6}.}

@@

Our analysis of B1555+375 demonstrates that the choice of the macromodel used to describe the `smooth' lens galaxy mass distribution, combined with the limited information provided by the position of compact lensed images, can lead to a misinterpretation of the origin of flux-ratio anomalies in some systems. Special care may be needed for small-separation lens systems, in which the lensing galaxy is likely to be a spiral \citep[e.g.,][]{Turner84, F92}.  This is because a disc with a high inclination angle can significantly affect the lensing cross-section for such systems \citep{Maller97,Kee98,Moller03}. The current sample of lensed quasars with radio-wavelength flux ratio measurements is small.  Therefore, having even a few systems with edge-on discs may be sufficient to introduce a bias into the substructure analysis. In fact, our near-infrared AO imaging of
other lenses in the SHARP sample shows that B1555+375 is not the only
system that has a significant disc component that crosses the merging lensed images (Fassnacht et al., in prep.).  
These gravitational lenses will be investigated in future
work, but the images alone suggest that other systems may have a similar
link between flux-ratio anomalies and disc components. This potential bias introduced by edge-on discs may be avoided by post-selection as  sample sizes grow in the future. Therefore, it is critical to obtain high-resolution imaging lenses to detect additional mass features such as edge-on discs or even massive luminous satellites.
% A number of lens systems with significant disc components arealready known, including, for example, the sample of 16 edge-on disks from SDSS found in the SWELLS survey \citep{swells1}. 
 This result is timely, as upcoming large-scale surveys (e.g., the Dark Energy Survey, Kilo-Degree Survey, \textit{Euclid}, and the Large Synoptic Survey Telescope) will deliver orders of magnitude more gravitationally lensed quasars for which flux ratios can
  be measured using narrow lines \citep[e.g.][]{N14}.
  %Finally, we note that massive edge-on discs are \textit{not} observed in some flux-ratio anomaly systems \citep[and the presence of a disc-like component may not resolve the anomaly anyway;][]{evans03}, indicating that these systems are still useful for probing the dark matter substructure mass function.
